---
title: "Programming Assignment ML"
author: "Jordi Overeem"
date: "22 juni 2018"
output: html_document
---

# Practical Machine Learning assignment
## Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


The data for this project come from this source: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

## Loading packages and the data
```{r}
# SETTINGS & LIBRARIES
setwd("C:/Users/joov2/Desktop")
set.seed(555)
library(caret)
library(randomForest)
library(rattle)
library(corrplot)
library(rpart)

# LOAD DATAFRAMES
training_data_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_data_url <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
train <- data.frame(read.csv(file=training_data_url))
test <- data.frame(read.csv(file=test_data_url))
dim(train)
dim(test)
```
## Cleaning and pre-processing the data
With the code below, some ID variables and variables and cases with many missing values are removed from the dataset. Also, the data is partitioned and prepared for validation.
```{r}
# DATA CLEANSING
train <- train[,-c(1:6)] # Remove ID vars
test <- test[,-c(1:6)]

NZs <- nearZeroVar(train)
train <- train[, -NZs] # Remove vars with many missing values
test <- test[, -NZs]

NAs <- sapply(train, function(x) mean(is.na(x))) > 0.95
train <- train[, NAs == "FALSE"] # Remove cases with many NA's
test <- test[, NAs == "FALSE"]
dim(train)
dim(test)
summary(train)
head(train)

# DATA PREPARATION
in_train = createDataPartition(train$classe, p = 3/4, list = FALSE)
train <- train[in_train, ]
crossval <- train[-in_train, ]
```

## Model 1 - Decision Tree model
First a decision tree model is trained.
```{r}
# MODEL 1 | DECISION TREE
mdl1 <- train(classe~., data = train, method = "rpart") # Decision tree model

predict_train1 <- predict(mdl1, train)
confmtrx_train1 <- confusionMatrix(predict_train1, train$classe) # Confusion matrix on training data

print(confmtrx_train1)

predict_crossval1 <- predict(mdl1, crossval)
confmtrx_crossval1 <- confusionMatrix(predict_crossval1, crossval$classe) # Confusion matrix on cross validation data

print(confmtrx_crossval1)
```
Accuracy of this model: 0.5736.

## Model 2 - Random Forest model
Second a random forest model is trained.
```{r}
# MODEL 2 | RANDOM FOREST
mdl2 <- randomForest(classe~., data = train, importance = TRUE, ntree = 10) # Random forest model
predict_train2 <- predict(mdl2, train)
confmtrx_train2 <- confusionMatrix(predict_train2, train$classe) # Confusion matrix on training data
varImpPlot(mdl2)
print(confmtrx_train2)

predict_crossval2 <- predict(mdl2, crossval)
confmtrx_crossval2 <- confusionMatrix(predict_crossval2, crossval$classe) # Confusion matrix on cross validation data

print(confmtrx_crossval2)
```
Accuracy of this model: 0.9997.

## Conclusion
Since the accuracy of the random forest model performs better than the decision tree model, the random forest model is used for predictions on the test data:
```{r}
predict <- predict(mdl2, test)
predict
```